---
title: "DATA 624 Homework 10"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

#### Name: Charles Ugiagbe.
#### Date: 12/16/23


###   Homework Intro 
The aim of this assigment is to use  is to use R to mine the data for association rules. You should report support, confidence and lift and your top 10 rules by lift.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(arules)
library(igraph)
```


### load the data

```{r}
receipt.df <- read.csv("GroceryDataSet.csv")
```


### Take a head look at the data
```{r}
head(receipt.df)
```

### Market Analysis
we need to first read in and explore the data by looking at the top 15 item purchase.

```{r}
data.df <- read.transactions("GroceryDataSet.csv", sep=",")
itemFrequencyPlot(data.df, topN=15, type="absolute", main="Top 15 Items", col=rainbow(15))
```
**Whole milk is the most frequently purchased item.**


### Association Rules

In order to complete this market basket analysis, the Apriori algorithm is initiated to print out the top 10 rules with their support, confidence and lift.
To find the association rules, we will use the ‘apriori’ function.

```{r}
rules<- apriori(data.df, parameter=list(supp=0.001, conf=0.5) , control=list(verbose=FALSE))
summary(rules)
```

```{r}
apriori(data.df, parameter=list(supp=0.001, conf=0.5) , control=list(verbose=FALSE)) %>%
  DATAFRAME() %>%
  arrange(desc(lift)) %>%
  top_n(10) %>%
  kable() %>%
  kable_styling()
```

### Cluster Analysis

We are to look for item groupings. A network graph can be use to preform the cluster analysis. First I will need to create a network graph from the transaction data. The I will detect the communities in the graph using the Louvain algorthym.

```{r}
temp <- read.csv("GroceryDataSet.csv", header = FALSE) %>%
  mutate(shoper_id = row_number()) %>%
  pivot_longer(-shoper_id) %>%
  filter(value != "") %>%
  select(-name)

louvain_communities <- temp %>%
  rename(to = value, from = shoper_id) %>%
  graph_from_data_frame(directed = FALSE) %>%
  cluster_louvain() %>%
  communities()
```


```{r}
items <- as.character(unique(temp$value))

cluster_df <- data.frame(name = c(NA), members = c(NA)) %>% na.omit()

for (i in 1:length(louvain_communities)){
  cluster_name <- paste0(i,": ")
  cluster_members <- 0
  for (member in louvain_communities[[i]]){
    if (member %in% items){
      cluster_name <- paste0(cluster_name, member, " + ")
    } else {
      cluster_members <- cluster_members + 1
    }
  }
  cluster_name <- substr(cluster_name,1,nchar(cluster_name)-3)
  cluster_df <- rbind(cluster_df, data.frame(name = cluster_name, members = cluster_members))
}

cluster_df %>%
  arrange(desc(members)) %>%
  kable() %>%
  kable_styling()
```



